Still Missing a Cool Title

### 2006-05-31 talk/dls2006/paper.tex
i.e. => i.e.\

\begin{abstract}
The PyPy project seeks to prove both on a research and a
practical level the feasibility of writing a virtual machine (VM)
for a dynamic language in a dynamic language -- in this case, Python.
The aim is to translate (i.e. compile) the VM to arbitrary target
environments, ranging in level from C/Posix to Smalltalk/Squeak via
Java and CLI/.NET, while still being of reasonable efficiency within
these environments.

@@

\item We perform type inference on the control flow graphs.  At this
      stage, types inferred are part of the type system which is the
      very definition of the RPython sub-language: they are roughly a
      subset of Python's built-in types, with some more precision to
      describe e.g.\ the items stored in container types.
      Occasionally, a single input function can produce several
      specialized versions, i.e. several similar but differently typed
      graphs.  This type inference process is described in more
      details in section \ref{typeinference}.
\end{enumerate}
 %     
At the end of the front-end analysis, the input RPython program is
represented as a forest of flow graphs with typed variables.  Following
this analysis are a number of transformation steps.  Each transformation
step modifies the graphs in-place, by altering their structure and/or
the operations they contain.  Each step inputs graphs typed in one type
system and leaves them typed in a possibly different type system, as we
will describe in the sequel.  Finally, a back-end turns the resulting
graphs into code suitable for the target environment, e.g.\ C source code
ready to be compiled.

@@

The actual transformation from function objects -- i.e. bytecode -- to
flow graph is performed by the Flow Object Space, a short but generic
plug-in component for the Python interpreter of PyPy.  The architecture
of our Python interpreter is shown in figure \ref{interpobjspace}.

@@

Note that the left column, i.e. the bytecode interpreter and the
Standard Object Space, form the full Python interpreter of PyPy.  It is
an RPython program, and the whole purpose of the translation process is
to accept this as \textit{input}, and translate it to an efficient form.  Its
architecture is not relevant to the way it is translated. XXX?

@@

More precisely, it is usually possible to deduce information about the
result variable of an operation given information about its arguments.
For example, we can say that the addition of two integers must be an
integer.  Most programming languages have this property.  However,
Python -- like many languages not specifically designed with type
inference in mind -- does not possess a type system that allows much
useful information to be derived about variables based on how they are
\textit{used}; only on how they were \textit{produced}.  For example, a number of very
different built-in types can be involved in an addition; the meaning of
the addition and the type of the result depends on the type of the input
arguments.  Merely knowing that a variable will be used in an addition
does not give much information per se.  For this reason, our annotator
works by flowing types forward, operation after operation, i.e. by
performing abstract interpretation of the flow graphs.  In a sense, it
is a more naive approach than the one taken by type systems specifically
designed to enable more advanced inference algorithms.  For example,
Hindley-Milner type inference works in an inside-out direction, by
starting from individual operations and propagating type constraints
outwards [H-M].

@@

To achieve high performance for dynamic languages such as Python, the
proven approach is to use dynamic compilation techniques, i.e. to write
JITs.  With direct techniques, this is however a major endeavour, and
increases the efforts to further evolve the language.

>>>>

\begin{abstract}
The PyPy project seeks to prove both on a research and a
practical level the feasibility of writing a virtual machine (VM)
for a dynamic language in a dynamic language -- in this case, Python.
The aim is to translate (i.e.\ compile) the VM to arbitrary target
environments, ranging in level from C/Posix to Smalltalk/Squeak via
Java and CLI/.NET, while still being of reasonable efficiency within
these environments.

@@

\item We perform type inference on the control flow graphs.  At this
      stage, types inferred are part of the type system which is the
      very definition of the RPython sub-language: they are roughly a
      subset of Python's built-in types, with some more precision to
      describe e.g.\ the items stored in container types.
      Occasionally, a single input function can produce several
      specialized versions, i.e.\ several similar but differently typed
      graphs.  This type inference process is described in more
      details in section \ref{typeinference}.
\end{enumerate}
 %     
At the end of the front-end analysis, the input RPython program is
represented as a forest of flow graphs with typed variables.  Following
this analysis are a number of transformation steps.  Each transformation
step modifies the graphs in-place, by altering their structure and/or
the operations they contain.  Each step inputs graphs typed in one type
system and leaves them typed in a possibly different type system, as we
will describe in the sequel.  Finally, a back-end turns the resulting
graphs into code suitable for the target environment, e.g.\ C source code
ready to be compiled.

@@

The actual transformation from function objects -- i.e.\ bytecode -- to
flow graph is performed by the Flow Object Space, a short but generic
plug-in component for the Python interpreter of PyPy.  The architecture
of our Python interpreter is shown in figure \ref{interpobjspace}.

@@

Note that the left column, i.e.\ the bytecode interpreter and the
Standard Object Space, form the full Python interpreter of PyPy.  It is
an RPython program, and the whole purpose of the translation process is
to accept this as \textit{input}, and translate it to an efficient form.  Its
architecture is not relevant to the way it is translated. XXX?

@@

More precisely, it is usually possible to deduce information about the
result variable of an operation given information about its arguments.
For example, we can say that the addition of two integers must be an
integer.  Most programming languages have this property.  However,
Python -- like many languages not specifically designed with type
inference in mind -- does not possess a type system that allows much
useful information to be derived about variables based on how they are
\textit{used}; only on how they were \textit{produced}.  For example, a number of very
different built-in types can be involved in an addition; the meaning of
the addition and the type of the result depends on the type of the input
arguments.  Merely knowing that a variable will be used in an addition
does not give much information per se.  For this reason, our annotator
works by flowing types forward, operation after operation, i.e.\ by
performing abstract interpretation of the flow graphs.  In a sense, it
is a more naive approach than the one taken by type systems specifically
designed to enable more advanced inference algorithms.  For example,
Hindley-Milner type inference works in an inside-out direction, by
starting from individual operations and propagating type constraints
outwards [H-M].

@@

To achieve high performance for dynamic languages such as Python, the
proven approach is to use dynamic compilation techniques, i.e.\ to write
JITs.  With direct techniques, this is however a major endeavour, and
increases the efforts to further evolve the language.
